---
title: "Copying_Databases"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Copying_Databases}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval=FALSE
)
```

# Copying Data Between Experiments

*Note: this vignette is not for those working with the graphical user interface, but for those wishing to work with IDBac data in R*

Sometimes it is desireable to copy data from a database to a new database or another existing one. 
IDBac contains a two functions to help you do this. 

If you haven't already, you will need to install the IDBacApp R package, this will also install the necessary pacakge dependencies. 

The first function is:
```{r}
conn <- IDBacApp::createPool(fileName, 
                             filePath)
```

This uses the pool package ([read more about pool](https://shiny.rstudio.com/articles/pool-basics.html)).  If there is a sqlite file that exists at `filePath`, with the name `filePath`, it will return a reference to it. If the file doesn't exist it will create a new sqlite file and return a reference to that new file. `IDBacApp::createPool` returns a list of pool objects. This means that you can pass multiple fileNames/filePaths, in equal length and createPool will return a list of equal length.



The second function is:

```{r}
copyToNewDatabase(existingDBPool,
                  newDBPool, 
                  sampleIDs)

```

This one is hopefully straightforward.

- `existingDBPool` is a single (not list) pool object referencing the database to transfer **from**
- `newDBPool` is a single (not list) pool object referencing the database to transfer **to**
- `sampleIDs` is a character vector of sampleIDs that exist in `existingDBPool` and you want to transfer to `newDBPool`



So, what this looks like in practice:


```{r}
# Using example data within the IDBac package:
filePath <- system.file("extdata/sqlite", package = "IDBacApp")

# Notice that we only include the file name, without the '.sqlite' extension
fileName <- "sample"

# Make the pool object
conn <- IDBacApp::createPool(fileName, 
                             filePath)

# We don't need the conn variable as a list, so we extract the pool object
conn <- conn[[1]]

```

We can inspect the database
```{r}
DBI::dbListTables(conn)
```


Let's take a peek at which samples are included in the sample database:
```{r}
DBI::dbGetQuery(conn, 
                "SELECT DISTINCT Strain_ID FROM metaData")
```

Create the new database to transfer to:
```{r}

tempDatabase <- tempfile()

connNew <- IDBacApp::createPool(fileName = basename(tempDatabase), 
                             dirname(tempDatabase))
connNew <- connNew[[1]]
```



```{r}
copyToNewDatabase(existingDBPool = conn,
                  newDBPool = connNew, 
                  sampleIDs = "sample")
```

